`infercnv` results exploration for `r params$sample_id`

### 2024-10-18 analyses/cell-type-wilms-tumor-06/notebook_template/06_cnv_infercnv_exploration.Rmd
changes to PR#828

---
title: "`infercnv` results exploration for `r params$sample_id`"
author: "Maud PLASCHKA"
date: "`r Sys.Date()`"
params:
  scpca_project_id: "SCPCP000006"
  sample_id: "SCPCS000194"
  seed: 12345
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

@@

library(infercnv)
library(SCpubr)
library(ggplot2)
library(patchwork)

@@

@@

Do_CNV_heatmap <- function(object, reference_value, group.by){

  infercnv_obj <- readRDS(file.path(result_dir, reference_value , glue::glue("06_infercnv_", params$sample_id, "_", reference_value, ".rds")))

@@

  out <- SCpubr::do_CopyNumberVariantPlot(sample = srat,
                                        infercnv_object = infercnv_obj,
                                        using_metacells = FALSE,
                                        chromosome_locations = chromosome_locations,
                                        return_object = FALSE,
                                        group.by = group.by
                                        )
  return(out)
}

@@

@@

- `metadata` the gene or quantitative value to be plotted

@@

  if(is.numeric(object@meta.data[,meta])){
     d <- SCpubr::do_FeaturePlot(object, 
                                              features = meta, 
                                              pt.size = 0.2, 
                                              legend.width = 0.5, 
                                              legend.length = 5, 
                                              legend.position = "right") + ggtitle(meta)
    b <- SCpubr::do_ViolinPlot(srat, 
                                             features = meta, 
                                             ncol = 1, 
                                             group.by = group.by, 
                                             legend.position = "none")

@@

                  d <- SCpubr::do_FeaturePlot(object, 
                                              features = features, 
                                              pt.size = 0.2, 
                                              legend.width = 0.5, 
                                              legend.length = 5, 
                                              legend.position = "right") + ggtitle(features)
                  b <- SCpubr::do_ViolinPlot(srat, 
                                             features = features, 
                                             ncol = 1, 
                                             group.by = group.by, 
                                             legend.position = "none",
                                             assay = "SCT") + ylab(features)

@@

reference_value = "reference-none"

@@

![](`r file.path(result_dir, reference_value , glue::glue("06_infercnv_", params$sample_id, "_", reference_value, "_heatmap.png"))`)

Do_CNV_heatmap(object = srat, reference_value, group.by = "fetal_kidney_predicted.compartment")

@@

reference_value = "reference-immune"

![](`r file.path(result_dir, reference_value , glue::glue("06_infercnv_", params$sample_id, "_", reference_value, "_heatmap.png"))`)

Do_CNV_heatmap(object = srat, reference_value, group.by = "fetal_kidney_predicted.compartment")

reference_value = "reference-endothelium"

![](`r file.path(result_dir, reference_value , glue::glue("06_infercnv_", params$sample_id, "_", reference_value, "_heatmap.png"))`)

Do_CNV_heatmap(object = srat, reference_value, group.by = "fetal_kidney_predicted.compartment")

reference_value = "reference-both"

![](`r file.path(result_dir, reference_value , glue::glue("06_infercnv_", params$sample_id, "_", reference_value, "_heatmap.png"))`)

Do_CNV_heatmap(object = srat, reference_value, group.by = "fetal_kidney_predicted.compartment")

@@

srat <- readRDS(file.path(module_base, "results", params$sample_id, glue::glue("06_infercnv_", params$sample_id, "_reference-both.rds")))

@@

###  Distribution of CNV estimation in the Wilms tumor copartments

@@

```{r fig.width=5, fig.height=5, out.width='40%', out.height='40%'}
for(i in 1:22){
  print(visualize_density(srat, features=glue::glue("proportion_cnv_chr", i), group.by = "fetal_kidney_predicted.compartment"))
}
```

@@

We should maybe try to run `infercnv` on one sample, gradualy downsampling the reference to 50, 10, 5, 1 cell to see how it impact the results. 
For this try, I would suggest running with the HMM i3 model, to have an additional idea of the feasibility to extract CNV features from `infercnv`.

>>>>

---
title: "`infercnv` results exploration for `r params$sample_id`"
author: "Maud PLASCHKA"
date: "`r Sys.Date()`"
params:
  scpca_project_id: "SCPCP000006"
  sample_id: "SCPCS000179"
  seed: 12345
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

@@

library(infercnv)
library(SCpubr)
library(ggplot2)
library(patchwork)
library(Seurat)

@@

- `infercnv_obj` is the `infercnv` object

@@

Do_CNV_heatmap <- function(object, infercnv_obj, group.by){

@@

  out <- SCpubr::do_CopyNumberVariantPlot(sample = object,
                                        infercnv_object = infercnv_obj,
                                        using_metacells = FALSE,
                                        chromosome_locations = chromosome_locations,
                                        return_object = FALSE,
                                        group.by = group.by
                                        )
  out <- out + 
    ggtitle(glue::glue("Copy Number Variant Plot, ", reference_value)) +
    ylab(label = "")

  return(out)
}

@@

#### Calculate a global CNV score per cell to check the general distribution

For a `Seurat` object an `infercnv` object created with the script `06_infercnv.R` using `reference_value` as a reference, the function `Do_CNV_score` load the `infercnv` object and calculate a CNV score per cell.
The score is calculated based on the [biostar discussion](https://www.biostars.org/p/9573777/).
It

Do_CNV_score <- function(seurat_oject,infercnv_obj,reference_value){

  score <- apply(infercnv_obj@expr.data,2,function(x){ sum(x < 0.90 | x > 1.10)/length(x) })

  seurat_obj <- AddMetaData(seurat_oject, score, col.name = glue::glue("CNV-score_", reference_value))
  return(seurat_obj)
}
```

@@

- `meta` the gene or quantitative value to be plotted

@@

  if(is.numeric(object@meta.data[,meta])){
     d <- SCpubr::do_FeaturePlot(object, 
                                              features = meta, 
                                              pt.size = 0.2, 
                                              legend.width = 0.5, 
                                              legend.length = 5, 
                                              legend.position = "right") + ggtitle(meta)
    b <- SCpubr::do_ViolinPlot(object, 
                                             features = meta, 
                                             ncol = 1, 
                                             group.by = group.by, 
                                             legend.position = "none")

@@

                  d <- SCpubr::do_FeaturePlot(object, 
                                              features = features, 
                                              pt.size = 0.2, 
                                              legend.width = 0.5, 
                                              legend.length = 5, 
                                              legend.position = "right") + ggtitle(features)
                  b <- SCpubr::do_ViolinPlot(object, 
                                             features = features, 
                                             ncol = 1, 
                                             group.by = group.by, 
                                             legend.position = "none",
                                             assay = "SCT") + ylab(features)

@@

![](`r file.path(result_dir, "reference-none_HMM-no" , glue::glue("06_infercnv_", params$sample_id, "_", "reference-none_HMM-no", "_heatmap.png"))`)

![](`r file.path(result_dir, "reference-immune_HMM-no" , glue::glue("06_infercnv_", params$sample_id, "_", "reference-immune_HMM-no", "_heatmap.png"))`)

![](`r file.path(result_dir, "reference-endothelium_HMM-no" , glue::glue("06_infercnv_", params$sample_id, "_", "reference-endothelium_HMM-no", "_heatmap.png"))`)

![](`r file.path(result_dir, "reference-both_HMM-no" , glue::glue("06_infercnv_", params$sample_id, "_", "reference-both_HMM-no", "_heatmap.png"))`)

### Summarize CNV per chromosome and compartment

```{r fig.width=20, fig.height=5, out.width='100%', out.height='100%'}
infercnv_obj <- list()
for(reference_value in c("reference-none", "reference-immune", "reference-endothelium", "reference-both")){
  infercnv_obj[[reference_value]] <- readRDS(file.path(result_dir, glue::glue(reference_value, "_HMM-no") , glue::glue("06_infercnv_", params$sample_id, "_", reference_value, "_HMM-no.rds")))
  print(Do_CNV_heatmap(object = srat, infercnv_obj = infercnv_obj[[reference_value]], group.by = "fetal_kidney_predicted.compartment"))
}

@@

### Try to calculate  single CNV score and see if it can be use to define cells with CNV

```{r fig.width=20, fig.height=5, out.width='100%', out.height='100%', warning=FALSE}

for(reference_value in c("reference-none", "reference-immune", "reference-endothelium", "reference-both")){
  srat <- Do_CNV_score(srat,infercnv_obj = infercnv_obj[[reference_value]], reference_value)
  p1 <- visualize_feature(srat, glue::glue("CNV-score_", reference_value) , group.by = "fetal_kidney_predicted.compartment" )
  p2 <- visualize_density(srat, features=glue::glue("CNV-score_", reference_value), group.by = "fetal_kidney_predicted.compartment")
  print(p1 + p2 + plot_layout(ncol = 3, widths = c(1,2,2)))
}

@@

This unique CNV score do not look so promising. 
We might have to select chromosomes we would like to look at, i.e. the one relevant for Wilms tumor.

### Explore infercnv results generated with immune and endothelial cells as reference, using a HMM i3 model

@@

srat <- readRDS(file.path(module_base, "results", params$sample_id, glue::glue("06_infercnv_HMM-i3_", params$sample_id, "_reference-both.rds")))

@@

####  Distribution of CNV estimation in the Wilms tumor copartments

@@

```{r fig.width=5, fig.height=5, out.width='40%', out.height='40%', warning=FALSE}
for(i in 1:22){
  print(visualize_density(srat, features=glue::glue("proportion_cnv_chr", i), group.by = "fetal_kidney_predicted.compartment"))
}
```

srat <- readRDS(file.path(module_base, "results", params$sample_id, glue::glue("06_infercnv_HMM-i6_", params$sample_id, "_reference-both.rds")))

####  Distribution of CNV estimation in the Wilms tumor copartments

```{r fig.width=5, fig.height=5, out.width='40%', out.height='40%', warning=FALSE}
for(i in 1:22){
  print(visualize_density(srat, features=glue::glue("proportion_cnv_chr", i), group.by = "fetal_kidney_predicted.compartment"))
}
```

@@

I am not sure if some endothelial cells might be missclassified? 
To be careful when we look at % of alterations within a compartment, as there are sometimes very few immune and/or endothelial cells, some % can be easily high without being meaningful.

HMM-i3 model looks "cleaner" than the HMM-i6 model, with less CNV found in immune and endothelial cells.
